---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# solvergater

<!-- badges: start -->
[![R build status](https://github.com/maciejsmolka/solvergater/workflows/R-CMD-check/badge.svg)](https://github.com/maciejsmolka/solvergater/actions)
[![Codecov test coverage](https://codecov.io/gh/maciejsmolka/solvergater/branch/master/graph/badge.svg)](https://codecov.io/gh/maciejsmolka/solvergater?branch=master)
<!-- badges: end -->

It is a common task to solve an optimization problem when the objective is computed 
using an external application such as a finite-element-method partial differential
equation solver that gets 
its input from the command line and writes its output to a number of files. 
The `solvergater` package provides a simple R gateway to such external apps.

## Installation

<!--
You can install the released version of solvergater from 
[CRAN](https://CRAN.R-project.org) with:
-->
<!--
``` r
install.packages("solvergater")
```
-->
<!--
And 
-->
You can install
the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("maciejsmolka/solvergater")
```
## Usage

Let us start with a basic example which shows you how to run an external solver to
compute the quantity of interest and its Jacobian matrix with respect to
the parameters. The example uses a mock solver provided with the package.

```{r exampleRun}
library(solvergater)
rscript_path <- file.path(R.home(), "bin", "Rscript")
solver_path <- file.path(find.package("solvergater"), "exec", "fake_simple.R")
nparams <- 4
nqoi <- 5
solver_cmd <- paste(rscript_path, solver_path, nparams, nqoi)
s <- shell_solver(solver_cmd, nparams = nparams, qoi_file = "output_qoi",
                  jacobian_file = "output_jacobian", wd = tempdir())
run(s, c(20.11, 5, -1.2, 10.4), silent = TRUE)
```

The basic use case for `solvergater` is the support of the optimization problems
with special attention paid for inverse problems. A simulated problem of the latter
type can be as follows.

First, we prepare some 'observed data' by running the solver at a high accuracy level.

```{r prepareData}
observed_data <- run(s, c(10, 10, 10, 10), silent = TRUE)$qoi
```

Then, as the objective we use the misfit between observed data and a current
proposed set of parameters. We can compute it in two ways. First, we can obtain
an objective function returning value and gradient at the same time.

```{r objective}
x <- c(10.5, 9.44, 10.21, 8.14)
obj <- objective(s, observed_data, silent = TRUE)
obj(x)
```

Second, we can obtain two separate functions for value and gradient. This form is
intended for the use in [`stats::optim()`](https://rdrr.io/r/stats/optim.html).

```{r optimization}
solver_obj <- objective_functions(s, observed_data, silent = TRUE)
solver_obj$value(x)
solver_obj$gradient(x)
```

`objective_functions()` [memoises](https://memoise.r-lib.org/) internal calls to an 
actual solver, so in the above example only a single solver run is performed.

```{r optimizationMemoise}
x1 <- c(0, 20, 0, 20)
solver_obj <- objective_functions(s, observed_data, silent = TRUE)
value_time <- system.time(solver_obj$value(x1))
gradient_time <- system.time(solver_obj$gradient(x1))
value_time
gradient_time
```
